---
title: "Randomized_results"
author: "Caitie"
date: "2026-05-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(tidyverse)
library(here)
library(terra)
```

# List files

```{r}
rand_overlap<-list.files(here("output_data/Random_turbines"), pattern = "overlap", full.names = TRUE)
rand_count<-list.files(here("output_data/Random_turbines"), pattern = "count", full.names = TRUE)
```

```{r}
rand_overlap_bind <- rand_overlap %>%
  map_dfr(~ read_csv(.x, col_types = cols(.default = col_character())) %>%
            select(-starts_with("...")) %>%
            mutate(source_file = basename(.x))) %>%
  mutate(
    across(
      c(feature_km2, rand_wind_km2, int_km2, prop_int, buff, iteration),
      as.numeric
    )
  )
```



```{r}
random_summary_overlap <- rand_overlap_bind %>%
  group_by(rgn_key, type, buff) %>%
  summarise(
    random_mean = mean(prop_int, na.rm = TRUE),
    random_median = median(prop_int, na.rm = TRUE),
    random_sd = sd(prop_int, na.rm = TRUE),
    q025 = quantile(prop_int, 0.025, na.rm = TRUE),
    q975 = quantile(prop_int, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

```


```{r}
test<-random_summary_overlap %>% 
  filter(random_mean>0)
```

# Convergence

```{r}
convergence <- rand_overlap_bind %>%
  arrange(rgn_key, type, buff, iteration) %>%
  group_by(rgn_key, type, buff) %>%
  mutate(
    running_mean = cummean(prop_int),
    running_median = sapply(seq_along(prop_int), function(i) median(prop_int[1:i])),
    running_q025 = sapply(seq_along(prop_int), function(i) quantile(prop_int[1:i], 0.025)),
    running_q975 = sapply(seq_along(prop_int), function(i) quantile(prop_int[1:i], 0.975))
  ) %>%
  ungroup()
```

```{r}
ggplot(convergence,
       aes(x = iteration, y = running_mean)) +
  geom_line() +
  facet_grid(type+buff ~ rgn_key, scales = "free_y") +
  theme_bw() +
  labs(
    x = "Number of random iterations",
    y = "Running mean overlap",
    title = "Convergence of random overlap estimates"
  )
```

```{r}
diagnostics <- rand_overlap_bind %>%
  group_by(rgn_key, type, buff) %>%
  summarise(
    n_iter = n_distinct(iteration),
    mean = mean(prop_int, na.rm = TRUE),
    sd = sd(prop_int, na.rm = TRUE),
    cv = ifelse(mean == 0, NA, sd / mean),
    .groups = "drop"
  )
```



# Read in actual results

```{r}
results<-read.csv(here("output_data/Paolo_results/Collated/Combined_area_by_country.csv")) %>% 
  select(rgn_key, feature_km2, buff, type, prop_wind_int) %>% 
  filter(type != "All")


```

```{r}
feature_iso3<-results %>% 
  select(rgn_key, feature_km2, type) %>% 
  distinct()
```


```{r}
join<-full_join(results, random_summary_overlap, by = c("rgn_key", "buff", "type"))

join2 <- join %>%
  mutate(across(where(is.numeric), ~replace_na(.x, 0)))

join3<-join2 %>% 
  filter(feature_km2>0)

```



```{r}
empirical_test <- rand_overlap_bind %>%
  group_by(rgn_key, buff, type, feature_km2) %>%
  summarise(
    random_values = list(prop_int[!is.na(prop_int)]),
    .groups = "drop"
  ) %>%
  left_join(results, by = c("rgn_key", "buff", "type", "feature_km2")) %>%
  mutate(prop_wind_int2 = replace_na(prop_wind_int, 0)) %>% 
  filter(prop_wind_int2>0) %>% 
  rowwise() %>%
  mutate(
      percentile = mean(unlist(random_values) < prop_wind_int2) +
        0.5 * mean(unlist(random_values) == prop_wind_int2),
      cat = case_when(
        percentile <= 0.05 ~ "Better than random",
        percentile >= 0.95 ~ "Worse than random",
        TRUE ~ "Random"
      )
    ) %>%
    ungroup()
```

```{r}
empirical_df<-as.data.frame(empirical_test) %>% 
  select(-random_values) %>% 
  filter(!is.na(feature_km2) == TRUE)

write.csv(empirical_df, here("output_data/Random_turbines/Random_empirical_results.csv"))

empirical_df25<-empirical_df %>% 
  filter(buff == 25)

write.csv(empirical_df25, here("output_data/Random_turbines/Random_empirical_results_25km.csv"))

empirical_df1<-empirical_df %>% 
  filter(buff == 1)

write.csv(empirical_df1, here("output_data/Random_turbines/Random_empirical_results_1km.csv"))

empirical_df15<-empirical_df %>% 
  filter(buff == 15)

write.csv(empirical_df15, here("output_data/Random_turbines/Random_empirical_results_15km.csv"))

```

# Test

```{r}
results_1km<-results %>% 
  filter(buff == 1)
```

At 1 km, 12 countries overlapped with a biodiversity features

CHL, RUS, FIN, DNK, DEU, POL, GBR, SWE, NLD, VNM, CHN, CAN

RUS, CHL, POL, CAN were removed due to not having enough turbines

# Countries omtited 

Var1 Freq
1  CAN    1
2  CHL    1
3  NOR    2
4  POL    3
5  RUS    2

# Missing countries

6 Ireland
7 Taiwan
8 South Korea
9 USA
10 France
11 Spain

# Included
12 Vietnam
13 United Kingdom
14 Sweden
15 Netherlands
16 Germany
17 Finland
18 Denmark
19 China



