---
title: "Untitled"
author: "Caitie"
date: "2025-03-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(here)
library(tidyverse)
```


EEZ file = eez_simple.shp
Infrastructure = sar_fixed_infrastructure_202411.csv

#Read in EEZ
```{r}
eez<-st_read(here("raw_data/EEZ/eez_simple.shp")) %>%
  st_make_valid(.)
```
```{r}
eez_diss<-eez %>%
  group_by(rgn_type) %>%
  summarise(.) %>%
  mutate(eez_area_km2 = as.numeric(st_area(.) *1e-6))
```

```{r}
eez_iso3_diss<-eez %>%
  group_by(rgn_key) %>%
  summarise(.) %>%
  mutate(eez_iso3_area_km2 = as.numeric(st_area(.) *1e-6))
```


# Dissolve infrastructure, calculate area

#Read in Infrastructure

```{r}
infrastructure_raw<-read_csv(here("raw_data/offshore_infrastructure/sar_fixed_infrastructure_202411.csv"))
```

# Let's keep only the useful columns (structure_id, lon, lat, label, label_confidence)

```{r}
infrastructure <- infrastructure_raw %>% 
  select(c(3,4,5,8,9))
```
 
# Keeping only the wind, oil and unknown categories + dissolving by structure_id

```{r}
infrastructure_positions <- infrastructure %>% 
  filter(label %in% c("wind", "oil", "unknown")) %>% 
  group_by(structure_id) %>% 
  summarise(
    lon = first(lon),
    lat = first(lat),
    label = first(label),
    label_confidence = first(label_confidence)) %>% 
  st_as_sf(coords = c("lon","lat"), crs = 4326)
```

# Reproject to same coordinate system as EEZ

```{r}
infrastructure_proj<-infrastructure_positions %>%
  st_transform(., crs = st_crs(eez))
```


# Add 50 km buffer to infrastructure

Units are in meters so 50 km = 50000 m

```{r}
infrastructure_buff<-st_buffer(infrastructure_proj, dist = 50000)
```


# Dissolve Infrastructure buffer, calculate area

Dissolve by label so we can keep oil, wind, unknown separate
If you wanted to dissolve them all together you would change what you group_by
```{r}
infrastructure_diss<-infrastructure_buff %>%
  group_by(label) %>%
  summarise(.) %>%
  mutate(infra_area_km2 = as.numeric(st_area(.) * 1e-6)) #calculate area and convert from m2 to km2
```

The area above is the total area of infrastructure given a 50 km buffer globally. We probably would want to know the difference between wind/oil/unknown so you would need

# Intersect EEZ with infrastructure within all EEZs and by country

```{r}
eez_infra<-st_intersection(eez_diss, infrastructure_diss) %>%
  mutate(eez_infra_km2 = as.numeric(st_area(.)*1e-6),
         prop_eez_infra = eez_infra_km2/eez_area_km2)
```

```{r}
eez_iso3_infra<-st_intersection(eez_iso3_diss, infrastructure_diss) %>%
  mutate(eez_iso3_infra_km2 = as.numeric(st_area(.)*1e-6),
         prop_eez_iso3_infra = eez_iso3_infra_km2/eez_iso3_area_km2)
```


